---
title: "STATS 201 Assignment 3"
author: 'Your Name and ID number here'
date: 'Due Date: ----'
output:
  word_document: default
  pdf_document: default
  html_document:
    fig_caption: yes
    number_sections: yes
---

```{r echo=FALSE,warning=FALSE}
## Do not delete this!
## It loads the s20x library for you. If you delete it 
## your document may not compile
require(s20x)
```

# Question 1 

## Question of interest/goal of the study

We wish to identify the most promising trombone practice method. Do any of the methods show significant improvement? If so, which method shows the most significant improvement and by how much?

## Read in and inspect the data:

```{r,fig.height=4.2,fig.width=6}
trombone.df <- read.csv("trombone.csv")
trombone.df$method <- factor(trombone.df$method,levels=c("NP","MP","MPS","PP","CP"))
stripchart(diff~method,data=trombone.df,vertical=T,pch=16,cex=1.5,method="stack",ylab="Difference in scores",main="Score change vs Training method")
summaryStats(diff~method,data=trombone.df)

```

## Comment on plot(s) and/or summary statistics:



 
## Fit model and check assumptions

```{r}



```



## Methods and assumption checks



## Executive Summary








****

# Question 2


## Question of interest/goal of the study



## Read in and inspect the data:

```{r,fig.height=5,fig.width=8}
height.df=read.table("height.txt",header=T)




```

## Comment on plot(s) and/or summary statistics:




## Fit model and check assumptions

```{r}


```

## Methods and assumption checks




## Executive Summary




